Mining Scenario/Transaction Pairs in HW/SW Co-verification Using Symbolic Execution

### 2017-04-06 paper.tex
minor

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                     Co-specification Model  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Lack of support for property specification in a HW/SW co-design is one of the
stumbling blocks for the application of formal techniques in co-verification.  
In this paper, we express a co-design property in C 
with the support of special primitive, called {\em next\_timeframe()}, 
that can express temporal correlation between HW and FW events. 
The co-design properties captures the overall design intent that enables
specification of system-level properties across HW/SW boundaries. Thus, 
they are sufficiently expressive, amenable to verification reuse and can be 
readily used in assertion-based verification.  Closest to our
unified property specification for co-verification is the language
\textit{xPSL}~\cite{xPSL} proposed by Xie et al., which is the extension of IEEE PSL.  
Note that the co-verification models in \hwcbmcv and \verifox are fully compliant 
with our specification language, hence easily verifiable.  
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %       Scenario, Transaction and Scenario/Transaction Pair
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A firmware only exercise some fragment of hardware state-space in a typical 
hardware/software
co-design~\cite{polig2014micro,polig2014fpl,giefers2015accelerating}.  
A firmware services the requests from the application software 
layer or its interacting hardware.  Similarly, the hardware services 
the requests from the firmware or the
physical environment.  Thus, the firmware and its interacting hardware
components can be seen as a set of services.  In this paper, we call a
hardware RTL service a {\em Hardware Transaction} $(\mathcal{HT})$.  Closest
to our notion of $\mathcal{HT}$ is the micro-architecture-level transaction
by Mahajan et al.~\cite{mcbmq}.  We call a FW service a {\em Firmware
Transaction} ($\mathcal{FT}$).  $\mathcal{FT}$ is an untimed finite-state
machine that is manifested as control-flow paths in the FW, similar to
transactions in TLM~\cite{codes14}.  A~{\em scenario} $(\mathcal{S})$ in a
co-design typically configures the co-verification environment in a way that
restricts the interaction between $\mathcal{FT}$ and $\mathcal{HT}$.  In
other words, given a scenario $\mathcal{S}$, the FW only exercises a fragment
of the HW state-space.  This renders many interactions between
$\mathcal{HT}$ and $\mathcal{FT}$ infeasible. For example, a scenario in 
\emph{Universal Asynchronous Receiver Transmitter} (UART) IP can configure the 
UART in \emph{transmitter} mode which restricts the communication between the FW 
and HW to set of services that only transfers data from the FW to the serial output 
port of the transmitter HW while completely bypassing the receiver HW.  
A {\em Scenario/Transaction} pair is a set of valid interactions between 
$\mathcal{HT}$ and $\mathcal{FT}$ for a particular scenario~$\mathcal{S}$.  
It is necessary to automatically mine Scenario/Transaction pairs for scalable 
and effective co-verification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %        Difference between Monolithic and Path-wise Approach
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A monolithic approach in \hwcbmcv is based on Bounded Model Checking~\cite{CKY03}, 
hence it cannot exploit the infeasibility arising from scenario driven
partitioning of HW-FW interaction behaviors.  Whereas, a path-based symbolic 
execution in \verifox can automatically infer such Scenario/Transaction 
pairs and check their validity using backend solvers.  We experimentally 
show that for building an effective and scalable formal co-verification tool, 
it is imperative to automatically mine only relevant scenario/transaction 
pairs that occurs in practical co-design environment. 
%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                     Co-specification Model  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Lack of support for property specification in a HW/SW co-design is one of the
stumbling blocks for the application of formal techniques in co-verification.  
In this paper, we express a co-design property in C 
with the support of special primitive, called {\em next\_timeframe()}, 
that can express temporal correlation between HW and FW events. 
The co-design properties captures the overall design intent that enables
specification of system-level properties across HW/SW boundaries. Thus, 
they are sufficiently expressive, amenable to verification reuse and can be 
readily used in assertion-based verification.  Closest to our
unified property specification for co-verification is the language
\textit{xPSL}~\cite{xPSL} proposed by Xie et al., which is the extension of IEEE PSL.  
Note that the co-verification models in \hwcbmcv and \verifox are fully compliant 
with our specification language, hence easily amenable to verification.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %       Scenario, Transaction and Scenario/Transaction Pair
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A firmware only exercises a fraction of the hardware state-space in a typical 
hardware/software
co-design~\cite{polig2014micro,polig2014fpl,giefers2015accelerating}.  
A firmware services the requests from the application software 
layer or its interacting hardware.  Similarly, the hardware services 
the requests from the firmware or the
physical environment.  Thus, the firmware and its interacting hardware
components can be seen as a set of services.  In this paper, we call a
hardware RTL service a {\em Hardware Transaction} $(\mathcal{HT})$.  Closest
to our notion of $\mathcal{HT}$ is the micro-architecture-level transaction
by Mahajan et al.~\cite{mcbmq}.  We call a FW service a {\em Firmware
Transaction} ($\mathcal{FT}$).  $\mathcal{FT}$ is an untimed finite-state
machine that is manifested as control-flow paths in the FW, similar to
transactions in TLM~\cite{codes14}.  A~{\em scenario} $(\mathcal{S})$ in a
co-design typically configures the co-verification environment in a way that
restricts the interaction between $\mathcal{FT}$ and $\mathcal{HT}$.  In
other words, given a scenario $\mathcal{S}$, the FW exercises only a fragment
of the HW state-space.  This renders many interactions between
$\mathcal{HT}$ and $\mathcal{FT}$ infeasible. For example, a scenario in 
\emph{Universal Asynchronous Receiver Transmitter} (UART) IP can configure the 
UART in \emph{transmitter} mode which restricts the communication between the FW 
and HW to set of services that only transfer data from the FW to the serial output 
port of the transmitter HW while completely bypassing the receiver HW.  
A {\em Scenario/Transaction} pair is a set of valid interactions between 
$\mathcal{HT}$ and $\mathcal{FT}$ for a particular scenario~$\mathcal{S}$.  
It is necessary to automatically mine Scenario/Transaction pairs for scalable 
and effective co-verification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %        Difference between Monolithic and Path-wise Approach
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A monolithic approach in \hwcbmcv is based on Bounded Model Checking~\cite{CKY03}, 
hence it cannot exploit the infeasibility arising from scenario driven
partitioning of HW-FW interaction behaviors.  Whereas, a path-based symbolic 
execution in \verifox can automatically infer such Scenario/Transaction 
pairs and check their validity using backend solvers.  We experimentally 
show that for building an effective and scalable formal co-verification tool, 
it is imperative to automatically mine only relevant scenario/transaction 
pairs that occur in practical co-design environment. 
%
